Embedded Meta-Learning: Toward more flexible deep-learning models

### 2019-05-15 writing/embedded_meta_learning.tex
Tweaks and cutting

\section{Embedded Meta-Learning (EML) architecture}
EML architectures consist of two parts: 
\vspace{-0.75em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior based on examples). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[c.f.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

\caption{Schematic of our general architecure. Blocks with solid edges denote deep networks with learnable parameters, dashed edges represent inputs, outputs, embeddings, etc., and $F$ is a deep network with parameters specified by $\mathcal{H}$. Note that there may be multiple input, output, and target decoders if the system has multiple input modalities, e.g. both language and vision. See appendix \ref{app_model_details} for details.} \label{architecture_fig}

@@

Specifically, we embed inputs, targets, and mappings into a shared representational space $Z$. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Model outputs are decoded from the representational space by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Target outputs are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$.\par

@@

However, the system is not limited to simply mapping inputs to outputs. The key advantage of our architecture is that it allows transforming anything that is embedded in $Z$ using the same system. For example, suppose we have some task embedding $z_{old} \in Z$, and a meta-mapping embedding $z_{\text{meta}} \in Z$. Then we can generate a new task embedding $\hat{z}_{new} \in Z$ as:  

@@

\begin{figure}
\centering
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/meta_results.png}
\caption{From examples of mapping.}
\label{poly_meta_map_results_examples}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/language_meta_results.png}
\caption{From language input.}
\label{poly_meta_map_results_language}
\end{subfigure}
\caption{The system generalizes to apply learned meta-mappings to new tasks, and to apply unlearned meta-mappings from seeing examples or systematic language. The solid line indicates optimal performance; the dashed line indicates chance performance (an untrained model). The system is able to execute meta-mappings either from examples (\subref{cards_meta_map_results_examples}) or from language cues (\subref{cards_meta_map_results_language}). Error-bars are bootstrap 95\%-CIs.} 
\label{poly_meta_map_results}
\end{figure}
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par

@@

Our next case study was in the setting of simple card games, where the agent receives a hand and has to decide on a bet. There is no action besides betting, and depending on the opponent's hand the agent either wins or loses the amount bet. This doesn't require long term planning, but does incorporate some aspects of reinforcement learning, namely stochastic feedback on only the action chosen. We considered five games that are simplified analogs of various real card games (see Appendix \ref{meth_data_cards}). We also considered several binary options that could be applied to the games, including trying to lose instead of trying to win, or switching which suit was more valuable. These are challenging manipulations, for instance trying to lose requires completely inverting a learned $Q$-function. \par

@@

\section{Related work}
Our work is an extrapolation from the rapidly-growing literature on meta-learning \citep[e.g.][]{Vinyals2016, Santoro2016, Finn2017a, Finn2018, Stadie2018}. It is also related to the literature on continual learning, or more generally tools for avoiding catastrophic interference based on changes to the architecture \citep[e.g.][]{Fernando2017, Rusu2016}, loss \citep[e.g.][]{Kirkpatrick2016, Zenke2017, Aljundi2019}, or external memory \citep[e.g.][]{Sprechmann2018}. Recent work has also begun to blur the separation between these approaches, for example by meta-learning in an online setting \citep{Finn2019}. Our work is specifically inspired by the algorithms that attempt to have the system learn to adapt to a new task via activations rather than weight updates, either from examples \citep[e.g.][]{Wang2016a, Duan2016}, or by conditioning on a task input \citep[e.g.][]{Borsa2019}. \par
Our architecture builds directly off of prior work on HyperNetworks \citep{Ha2016} -- networks which parameterize other networks -- and other recent applications thereof, such as guessing parameters for a model to accelerate model search \citep[e.g.][]{Brock2018a, Zhang2019}. This is related to the longer history of work on different time-scales of weight adaptation \citep{Hinton1982, Kumaran2016} that has more recently been applied to meta-learning contexts \citep[e.g.][]{Ba2016, Munkhdalai2017} and continual learning \citep[e.g.]{Hu2019}. It is more abstractly related to work on learning to propose architectures \citep[e.g.][]{Zoph2016, Cao2019}, and to models that learn to select and compose skills to apply to new tasks \citep[e.g.][]{Andreas, Andreas2016, Tessler2016, Reed2015}. In particular, some of the work in domains like visual question answering has explicitly explored the idea of building a classifier conditioned on a question \citep{Andreas, Andreasa}, which is related to one of the possible computational paths through our architecture. Work in model-based reinforcement learning has also partly addressed how to transfer knowledge between different reward functions \citep[e.g.][]{Laroche2017}; our approach is more general. \par 

@@

\section{Future directions}
We think that the general perspective of embedding tasks and considering meta-mappings will yield many fruitful future directions. We hope that our work will inspire more exploration of behavioral adaptation in areas beyond the simple domains we considered here. To this end, we suggest the creation of meta-learning datasets which include information not only about tasks, but about the relationships between them. For example, as we alluded to above, tasks such as visual question-answering \citep[e.g.][]{Antol2015} can be usefully interpreted from this perspective (a question essentially yields a function from images to answers, and there are many systematic relationships between these functions), as can reinforcement learning tasks which involve executing instructions \citep[e.g.][]{Hermann2017, Co-Reyes2019}. Furthermore, we think our work provides a novel perspective on the types of flexibility that human intelligence exhibits, and thus hope that it may have implications for cognitive science. \par 

@@

\section{Conclusions}
In this paper, we've highlighted a new type of flexibility in the form of \emph{meta-mappings}, that is, mappings between tasks. We've suggested that achieving this type of flexibility is difficult in most deep learning architectures, because they separate computation from the knowledge of what is being computed. We've proposed that representing the task explicitly as the computation provides a solution to this problem. To this end, we've proposed a meta-learning system which embeds tasks and data into a shared latent space, and then meta-learns to make mappings over this space. The system can do this with quite small meta-task sample complexity, since it can generalize to unseen meta-mappings after training on only 20 examples.\par

@@

Note that there are two kinds of hold-out in the training of this system. First, some of the trained tasks are held-out on each training step (i.e. $D_2$), in order to force the system to generalize rather than just memorize. Second, there are truly held-out tasks that the system has never seen in training, these are the held-out tasks and meta-mappings that we evaluate on at the end of training.

>>>>

\section{Embedded Meta-Learning (EML) architecture}
EML architectures are composed of two components: 
\vspace{-0.75em}
\begin{enumerate} \setlength \itemsep{0em}
\item An input/output system: domain specific encoders and decoders (vision, language, etc.) that map into a shared embedding space $Z$.
\item A meta-learning system which embeds tasks into the shared embedding space $Z$, and learns to map task embeddings to behavior.
\end{enumerate}
\vspace{-0.75em}
The utility of having a completely shared space $Z$ in which data, language, and tasks are represented is that it allows for arbitrary mappings between these distinct types of data. In addition to basic tasks, the system could in principle learn to map language to tasks (follow instructions) or tasks to language (explaining behavior), or tasks to tasks (changing behavior based on examples). This is a step closer to the flexible reasoning humans display. \par
Without training on these mappings, of course, the system will not be able to execute them well. However, if it is trained on a broad enough set of such mappings, it will be able to generalize to new instances drawn from the same data/task distribution, as we demonstrate below. For instances that fall outside its data distribution, or for optimal performance, it may require some retraining, however. This reflects the structure of human behavior -- we are able to learn rapidly when new knowledge is relatively consistent with our previous knowledge, but learning an entirely new paradigm (such as calculus for a new student) can be quite slow \citep[c.f.][]{Kumaran2016}. \par 
\begin{figure}
\centering
\begin{tikzpicture}[auto, scale=0.75]
\node [dashblock] at (0, 0) (rep) {\begin{tabular}{c}$\bm{Z}$: \textbf{Shared} \\ \textbf{representation}\end{tabular}};

@@

\caption{The EML architecture. Blocks with solid edges denote learnable deep networks, dashed edges represent inputs, outputs, embeddings, etc., and $F$ is a deep network with parameters specified by $\mathcal{H}$. Note that there may be multiple input, output, and target decoders if the system has multiple input modalities, e.g. both language and vision. See appendix \ref{app_model_details} for details.} \label{architecture_fig}

@@

Specifically, we embed inputs, targets, and mappings into a shared representational space $Z$. Inputs are embedded by a deep network $\mathcal{I}: \text{input} \rightarrow Z$. Model outputs are decoded from $Z$ by a deep network $\mathcal{O}: Z \rightarrow \text{output}$. Target outputs are encoded by a deep network $\mathcal{T}: \text{targets} \rightarrow Z$.\par

@@

However, the system is not limited to simply mapping inputs to outputs. The key advantage of our architecture is that it allows transforming anything that is embedded in $Z$ using the same system. For example, suppose we have some task embedding $z_{old} \in Z$, and a meta-mapping embedding $z_{\text{meta}} \in Z$, e.g. from $z_{\text{meta}} = \mathcal{M}\left( \left\{\left((z_{source,0},z_{target,0}\right) \right\}\right)$. Then we can generate a new task embedding $\hat{z}_{new} \in Z$ as:  

@@

\begin{figure}
\centering
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/meta_results.png}
\caption{From examples of mapping.}
\label{poly_meta_map_results_examples}
\end{subfigure}%
\begin{subfigure}{0.5\textwidth}
\includegraphics[width=\textwidth]{figures/poly/language_meta_results.png}
\caption{From language input.}
\label{poly_meta_map_results_language}
\end{subfigure}
\caption{The system generalizes to apply learned meta-mappings to new tasks, and to apply unlearned meta-mappings, either from examples (\subref{poly_meta_map_results_examples}) or from language cues (\subref{poly_meta_map_results_language}). The solid line indicates optimal performance; the dashed line indicates chance (untrained). Error-bars are bootstrap 95\%-CIs.} 
\label{poly_meta_map_results}
\end{figure}
As a proof of concept, we first evaluated the system on the task of learning polynomials of degree $\leq 2$ in 4 variables (i.e. the task was to regress functions of the form $p: \mathbb{R}^4 \rightarrow \mathbb{R}$ where $p \in \mathcal{P}_2 \left(\mathbb{R}\right)$, though the model was given no prior inductive bias toward polynomial forms). This yields an infinite family of base-level tasks (the vector space of all such polynomials), as well as many families of meta-mappings over tasks (for example, multiplying polynomials by a constant, squaring them, or permuting their input variables). This allows us to not only examine the ability of the system to learn to learn polynomials from data, but also to adapt its learned representations in accordance with these meta-tasks. Details of the architecture and training can be found in appendix \ref{app_detailed_methods}.\par

@@

We next explored the setting of simple card games, where the agent is dealt a hand and must bet. There is no action besides betting, and depending on the opponent's hand the agent either wins or loses the amount bet. This doesn't require long term planning, but does incorporate some aspects of reinforcement learning, namely stochastic feedback on only the action chosen. We considered five games that are simplified analogs of various real card games (see Appendix \ref{meth_data_cards}). We also considered several binary options that could be applied to the games, including trying to lose instead of trying to win, or switching which suit was more valuable. These are challenging manipulations, for instance trying to lose requires completely inverting a learned $Q$-function. \par

@@

\section{Discussion}
\textbf{Related work:} Our work is an extrapolation from the rapidly-growing literature on meta-learning \citep[e.g.][]{Vinyals2016, Santoro2016, Finn2017a, Finn2018, Stadie2018}. It is also related to the literature on continual learning, or more generally tools for avoiding catastrophic interference based on changes to the architecture \citep[e.g.][]{Fernando2017, Rusu2016}, loss \citep[e.g.][]{Kirkpatrick2016, Zenke2017, Aljundi2019}, or external memory \citep[e.g.][]{Sprechmann2018}. Recent work has also begun to blur the separation between these approaches, for example by meta-learning in an online setting \citep{Finn2019}. Our work is specifically inspired by the algorithms that attempt to have the system learn to adapt to a new task via activations rather than weight updates, either from examples \citep[e.g.][]{Wang2016a, Duan2016}, or a task input \citep[e.g.][]{Borsa2019}. \par
Our architecture builds directly off of prior work on HyperNetworks \citep{Ha2016} -- networks which parameterize other networks -- and other recent applications thereof, such as guessing parameters for a model to accelerate model search \citep[e.g.][]{Brock2018a, Zhang2019}. This is related to the longer history of work on different time-scales of weight adaptation \citep{Hinton1982, Kumaran2016} that has more recently been applied to meta-learning contexts \citep[e.g.][]{Ba2016, Munkhdalai2017} and continual learning \citep[e.g.]{Hu2019}. It is more abstractly related to work on learning to propose architectures \citep[e.g.][]{Zoph2016, Cao2019}, and to models that learn to select and compose skills to apply to new tasks \citep[e.g.][]{Andreas, Andreas2016, Tessler2016, Reed2015, Chang2019a}. In particular, some of the work in domains like visual question answering has explicitly explored the idea of building a classifier conditioned on a question \citep{Andreas, Andreasa}, which is related to one of the possible computational paths through our architecture. Work in model-based reinforcement learning has also partly addressed how to transfer knowledge between different reward functions \citep[e.g.][]{Laroche2017}; our approach is more general. \par 

@@

\textbf{Future Directions:} We think that the general perspective of embedding tasks and considering meta-mappings will yield many fruitful future directions. We hope that our work will inspire more exploration of behavioral adaptation in areas beyond the simple domains we considered here. To this end, we suggest the creation of meta-learning datasets which include information not only about tasks, but about the relationships between them. For example, as we alluded to above, tasks such as visual question-answering \citep[e.g.][]{Antol2015} can be usefully interpreted from this perspective (a question essentially yields a function from images to answers, and there are many systematic relationships between these functions), as can reinforcement learning tasks which involve executing instructions \citep[e.g.][]{Hermann2017, Co-Reyes2019}. Furthermore, we think our work provides a novel perspective on the types of flexibility that human intelligence exhibits, and thus hope that it may have implications for cognitive science. \par 

@@

\textbf{Conclusions:} In this paper, we've highlighted a new type of flexibility in the form of \emph{meta-mappings}, that is, mappings between tasks. We've suggested that achieving this type of flexibility is difficult in most deep learning architectures, because they separate computation from the knowledge of what is being computed. We've proposed that representing the task explicitly as the computation provides a solution to this problem. To this end, we've proposed a meta-learning system which embeds tasks and data into a shared latent space, and then meta-learns to make mappings over this space. The system can do this with quite small meta-task sample complexity, since it can generalize to unseen meta-mappings after training on only 20 examples.\par

@@

Note that there are two kinds of hold-out in the training of this system. First, some of the trained tasks are held-out on each training step (i.e. $D_2$), in order to force the system to generalize rather than just memorize. Second, there are truly held-out tasks that the system has never seen in training, these are the held-out tasks and meta-mappings that we evaluate on at the end of training. \par
\textbf{Language-cued meta-tasks:} The procedure is analogous to the meta-tasks from examples, except that the input to $\mathcal{H}$ is the embedding of the language input, rather than the output of $\mathcal{M}$. \par
